---
title: "Homework 3"
author: "Sunny Kim"
date: "2025-03-03"
output: html_document
---
#Open file from downloads: 
```{r}
f <- "~/Desktop/zombies.csv" 
d <- read.table(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
```
## 1. Calculate population mean and standard deviation for each quantitative random variable (height, weight, age, number of zombies killed, and years of education)
```{r}
# Select quantitative variables
quant_vars <- d[, c("height", "weight", "age", "zombies_killed", "years_of_education")]
# Calculate population means
pop_means <- colMeans(quant_vars, na.rm = TRUE)
# Define population variance function
pop_var <- function(x) {mean((x - mean(x, na.rm = TRUE))^2, na.rm = TRUE)}
# Define population standard deviation function
pop_sd <- function(x){sqrt(pop_var(x))}
# Calculate population standard deviations
pop_sds <- sapply(quant_vars, pop_sd)
# Print results
pop_means
pop_sds
```
## 2. Use ggplot to make boxplots of quantitative variables by gender
```{r}
library(ggplot2)
#plot for gender and height 
p1 <- ggplot(data=d, aes(x = gender, y = height, color = gender))  #Define variables
p1 <- p1 + geom_boxplot()  #Graph them in boxplot
p1 <- p1 + xlab("Gender") + ylab("Height (cm)")  # Rename axes
p1 <- p1 + ggtitle("Zombie Height by Gender")  # Add title
p1
#plot for gender and weight
p2 <- ggplot(data=d, aes (x = gender, y = weight, color = gender))
p2 <- p2 + geom_boxplot()
p2 <- p2 + xlab("Gender") + ylab("Weight (kg)") 
p2 <- p2 + ggtitle("Zombie Weight by Gender")
p2  
#plot for gender and age
p3 <- ggplot(data = d, aes(x = gender, y = age, color = gender))  #
p3 <- p3 + geom_boxplot() 
p3 <- p3 + xlab("Gender") + ylab("Age (years)")  
p3 <- p3 + ggtitle("Zombie Age by Gender") 
p3  
#plot for gender and zombies killed
p4 <- ggplot(data = d, aes(x = gender, y = zombies_killed, color = gender))
p4 <- p4 + geom_boxplot()
p4 <- p4 + xlab("Gender") + ylab("Zombies Killed")
p4 <- p4 + ggtitle("Zombies Killed by Gender")
p4
#plot for gender and years of education 
p5 <- ggplot(data = d, aes(x = gender, y = years_of_education, color = gender))
p5 <- p5 + geom_boxplot()
p5 <- p5 + xlab("Gender") + ylab("Years of Education")
p5 <- p5 + ggtitle("Zombie Years of Education by Gender")
p5
```

## 3-1. Use ggplot to make scatterplots of height and weight in relation to age.
```{r}
#plot for height by age
p1 <- ggplot(data = d, aes(x = height, y = age)) #define variables
p1 <- p1 + geom_point() #create plot
p1 <- p1 + geom_smooth(method = "lm") #linear regression line
p1 <- p1 + ggtitle("Height by Age") + xlab("Age") + ylab("Height") #title + axes labels
p1
#plot for weight by age
p2 <- ggplot(data = d, aes(x = weight, y = age))
p2 <- p2 + geom_point()
p2 <- p2 + geom_smooth(method = "lm")
p2 <- p2 + ggtitle("Weight by Age") + xlab("Age") + ylab("Weight") 
p2
```
## 3-2. Do these variables seem to be related? In what way?
#Both show a positive linear trend which suggests that both height and weight increase with age

## 4-1. Using histograms and QQ plots, check whether the quantitative variables seem to be drawn from a normal distribution. 
```{r}
attach(d)
#Using histograms to check
par(mfrow = c(2, 3))  # Arrange plots in a 2-row, 3-column grid
# Height
hist(height, freq = FALSE, col = "skyblue", main = "Height Distribution",
     xlab = "Height", ylab = "Density")
lines(density(height, na.rm = TRUE), col = "black")
# Weight
hist(weight, freq = FALSE, col = "orange", main = "Weight Distribution",
     xlab = "Weight", ylab = "Density")
lines(density(weight, na.rm = TRUE), col = "black")

# Age
hist(age, freq = FALSE, col = "lightcoral", main = "Age Distribution",
     xlab = "Age", ylab = "Density")
lines(density(age, na.rm = TRUE), col = "black")

# Zombies killed
hist(zombies_killed, freq = FALSE, col = "lightyellow", main = "Zombie Kills Distribution",
     xlab = "Zombies Killed", ylab = "Density")
lines(density(zombies_killed, na.rm = TRUE), col = "black")

# Years of education
hist(years_of_education, freq = FALSE, col = "lightpink", main = "Years of Education Distribution",
     xlab = "Years of Education", ylab = "Density")
lines(density(years_of_education, na.rm = TRUE), col = "black")

#Using QQ plots to check
par(mfrow = c(2, 3))  # Arrange plots in a 2-row, 3-column grid
#Height
qqnorm(height, main = "Height Q-Q Plot") 
qqline(height, col = "red") 
#Weight
qqnorm(weight, main = "Weight Q-Q Plot")
qqline(weight, col = "red")
#Age 
qqnorm(age, main = "Age Q-Q Plot")
qqline(age, col = "red")
#Zombies killed
qqnorm(zombies_killed, main = "Zombies Killed Q-Q Plot")
qqline(zombies_killed, col = "red")
#ears of education 
qqnorm(years_of_education, main = "Years of Education Q-Q Plot")
qqline(years_of_education, col = "red")

par(mfrow = c(1, 1))  # Reset plotting layout for future plots
detach(d)
```
## 4-2. Which seem to be and which do not? For those that are not normal, can you determine from which common distribution they are drawn?
#Only height, weight and age seem to be drawn from a normal distribution. Guesses for rest: zombies killed - negative bionomial distribution, years of education - log-normal distribution 

## 5. Sample one subset of 30 zombie survivors (without replacement) and calculate mean and standard deviation for each variable. Also estimate standard error and construct the 95% confidence interval for each mean
```{r}
set.seed(30)
sample_size <- 30
#Sample from each variable
sample_height <- sample(height, sample_size, replace = FALSE)
sample_weight <- sample(weight, sample_size, replace = FALSE)
sample_age <- sample(age, sample_size, replace = FALSE)
sample_zombies_killed <- sample(zombies_killed, sample_size, replace = FALSE)
sample_years_of_education <- sample(years_of_education, sample_size, replace = FALSE)

library(dplyr) #load library

#Calculate mean and sd of sample variables
Mean_Height = mean(sample_height, na.rm = TRUE) #mean height
SD_Height = sd(sample_height, na.rm = TRUE) #sd height
SE_Height<- SD_Height/sqrt(sample_size) #se height

Mean_Weight = mean(sample_weight, na.rm = TRUE) #mean weight
SD_Weight = sd(sample_weight, na.rm = TRUE) #sd weight
SE_Weight<- SD_Weight/sqrt(sample_size) #se weight

Mean_Age = mean(sample_age, na.rm = TRUE) #mean age
SD_Age = sd(age, na.rm = TRUE) #sd age
SE_Age<- SD_Age/sqrt(sample_size) #se age

Mean_Zombies_killed = mean(zombies_killed, na.rm = TRUE) #mean zombies killed
SD_Zombies_killed = sd(zombies_killed, na.rm = TRUE) #sd zombies killed
SE_Zombies_killed<- SD_Zombies_killed/sqrt(sample_size) #se zombies killed

Mean_Education = mean(years_of_education, na.rm = TRUE) #mean years of education
SD_Education = sd(years_of_education, na.rm = TRUE) #sd years of education
SE_Education<- SD_Education/sqrt(sample_size) #se years of education

# Function for normal distribution 95% Confidence Intervals
normal_ci <- function(mean, se) {
  normal_lower <- mean - qnorm(1 - 0.05/2) * se
  normal_upper <- mean + qnorm(1 - 0.05/2) * se
  return(c(normal_lower, normal_upper))}
# Function for not normal distribution 95% Confidence Intervals
non_normal_ci <- function(mean, se, sample_size) {
  df <- sample_size - 1
  t_critical <- qt(1 - 0.05 / 2, df)
  lower <- mean - t_critical * se
  upper <- mean + t_critical * se
  return(c(lower, upper))}

#Calculate CIs
ci_height <- normal_ci(Mean_Height, SE_Height)
ci_weight <- normal_ci(Mean_Weight, SE_Weight)
ci_age <- normal_ci(Mean_Age, SE_Age)
ci_zombies_killed <- non_normal_ci(Mean_Zombies_killed, SE_Zombies_killed, sample_size)
ci_education <- non_normal_ci(Mean_Education, SE_Education, sample_size)

#create a data frame to store results
summary_table <- data.frame(
  Variable = c("Height", "Weight", "Age", "Zombies Killed", "Years of Education"),
  Mean = c(Mean_Height,Mean_Weight,Mean_Age,Mean_Zombies_killed,Mean_Education),
  SD = c(SD_Height,SD_Weight,SD_Age,SD_Zombies_killed,SD_Education),
  SE = c(SE_Height,SE_Weight,SE_Age,SE_Zombies_killed,SE_Education),
  CI_Lower = c(ci_height[1], ci_weight[1], ci_age[1], ci_zombies_killed[1], ci_education[1]),
  CI_Upper = c(ci_height[2], ci_weight[2], ci_age[2], ci_zombies_killed[2], ci_education[2]))
# Print the table with formatting
print(summary_table, row.names = FALSE, digits = 4)
```
## 6-1. Draw 99 more samples (total 100) and calculate means for each of these samples. 
```{r}
set.seed(30)
sample_size <- 30

library(dplyr) # load library
attach(d)

# Function to draw a sample and calculate means
draw_sample_means <- function() {
  sample_height <- sample(height, sample_size, replace = FALSE)
  sample_weight <- sample(weight, sample_size, replace = FALSE)
  sample_age <- sample(age, sample_size, replace = FALSE)
  sample_zombies_killed <- sample(zombies_killed, sample_size, replace = FALSE)
  sample_years_of_education <- sample(years_of_education, sample_size, replace = FALSE)
  return(data.frame(
    Height = mean(sample_height, na.rm = TRUE),
    Weight = mean(sample_weight, na.rm = TRUE),
    Age = mean(sample_age, na.rm = TRUE),
    Zombies_Killed = mean(sample_zombies_killed, na.rm = TRUE),
    Years_of_Education = mean(sample_years_of_education, na.rm = TRUE)))}
# Generate 99 new samples to make the total 100 samples
sampling_distribution <- bind_rows(draw_sample_means(), replicate(99, draw_sample_means(), simplify = FALSE))
#Print means for samples
print(sampling_distribution)
```
## 6-2. What are the means and sds of this distribution of means for each variable?
```{r}
# Calculate means and standard deviations for each variable in the sampling distribution
mean_sds <- data.frame(
  Variable = colnames(sampling_distribution),  
  Mean = sapply(sampling_distribution, mean),  
  SD = sapply(sampling_distribution, sd))
#Print results
print(mean_sds)
detach(d)
```
## 6-3. How do the standard deviations of means compare to the standard errors estimated in question 5?
#The standard deviations of means and standard errors from question 5 both measure variability but in different context. The SE refelcts the expected variation of a sample mean from the population mean whereas the SD reflects the variability across 100 sample means (aka sampling SD of sampling distribution).

## 6-4. What do these sampling distributions look like (graph)? Are they normally distributed? What about for those variables that you concluded were not originally drawn from a normal distribution?
```{r}
# Plot the sampling distributions of means for each variable using hist()
# Height
hist(sampling_distribution$Height, 
     main = "Sampling Distribution of Means: Height", 
     xlab = "Height", 
     col = "skyblue", 
     border = "black")
# Weight
hist(sampling_distribution$Weight, 
     main = "Sampling Distribution of Means: Weight", 
     xlab = "Weight", 
     col = "orange", 
     border = "black")
# Age
hist(sampling_distribution$Age, 
     main = "Sampling Distribution of Means: Age", 
     xlab = "Age", 
     col = "lightcoral", 
     border = "black")
# Zombies Killed
hist(sampling_distribution$Zombies_Killed, 
     main = "Sampling Distribution of Means: Zombies Killed", 
     xlab = "Zombies Killed", 
     col = "lightyellow", 
     border = "black")
# Years of Education
hist(sampling_distribution$Years_of_Education, 
     main = "Sampling Distribution of Means: Years of Education", 
     xlab = "Years of Education", 
     col = "lightpink", 
     border = "black")
```
#The variables that were originally normally distributed (height, weight, age) all have sampling distributions that appear to be noraml. The variables that were not originally drawn from normal distribution (zombies killed, years of education) also seem to have sampling distributions that appear to be fairly normal (more so compared to original distributions). This is a result of the Central Limit Theorem which applies even if the underlying data distribution is not normal.